---
title: "midterm-1-practice"
author: "elliothershberg"
date: "2022-04-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Practice problems for midterm 1.

## Extra problem sheet

```{r prob-1}
price <- c(5, 10, 15, 18, 22, 50, 100, 500)

median(price)

mean(price)

sd(price)
```

```{r prob-2}
# Can't calculate rate ratio or risk ratio, can calulcate odds ratio
smoker_odds <- (15 / 35) / (10 / 40)

smoker_odds
```

```{r prob-3}
smoker_rate <- 20 / (2 * 200)

# multiply by 100 to get in 100-person years
smoker_rate * 100

non_smoker_rate <- 50 / (3 * 800)

non_smoker_rate * 100

# rate ratio
smoker_rate / non_smoker_rate

# risk ratio = ratio of cumulative risks

# 20 out of 200 smokers had breast cancer
smoker_cum_risk <- 20 / 200

# 50 out of 800 non-smokers had breast cancer
non_smoker_cum_risk <- 50 / 800

# risk ratio
smoker_cum_risk / non_smoker_cum_risk

# odds ratio = the ratio of the odds of getting the disease
# odds = p / 1 - p
# in this case n with event / n - n with event
smoker_odds <- 20 / (200 - 20)
non_smoker_odds <- 50 / (800 - 50)

# odds ratio
smoker_odds / non_smoker_odds
```

There's an important concept here.
The relationship between the risk ratio and the odds ratio.
Odds ratios and risk ratios are similar for rare outcomes.
For common outcomes, odds ratios can be much larger, and that can be misleading.

```{r prob-4}
(4 / 310) * 1000

```

```{r prob-5}
(20 / 200) * (19 / 199)
```

6.

Conditional probabilities.
Prob pregnant, given a positive test.

P(pregnant | test+) = (P(test+ | pregnant) * P(pregnant)) / P(test+)

P(test+) = P(true positive) * P(pregnant) + P(false positive) * P(not pregnant)

```{r prob-6}
# If 50% of women taking test are pregnant
(0.95 * 0.5) / ( (0.95 * 0.5) + (0.02 * 0.5) )

# If 10% of women taking test are pregnant
(0.95 * 0.1) / ( (0.95 * 0.1) + (0.02 * 0.9) )
```

## Midterm prep questions

```{r prep-1}
0.168 * 0.168
```

```{r prep-2}
s_rate <- 25 / (2.1 * 200)

rate * 1000

n_rate <- 10 / (2.5 * 200)

s_rate / n_rate

s_risk <- 25 / 200
n_risk <- 10 / 200

s_risk / n_risk
```







