---
title: "r-distributions"
author: "elliothershberg"
date: "2022-05-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Some misc notes on using the functions for probability distributions in R.

## Function types

For any distribution, there are five function types:

- d: The probability density function
- p: The probability distribution function
- q: the quantile function
- r: random generation from the distribution

For example, dnorm is the density function for the normal distribution.

## Probability distribution usage

the pnorm function can return the area associated with a Z score.

```{r p-dist-usage}
# Z score of -1.66
pnorm(-1.66)
```

## Quantile usage

Opposite to the distribution, the quantile function can give a Z score for a certain area.

```{r}
# 25th percentile
qnorm(1 - 0.25)
```


## Binomial function

Determining probability of 0 out of 4 trials with 0.25 probability each:

```{r binom-examples}
dbinom(0, size = 4, prob = 0.25)
```



