---
title: "unit-9"
author: "elliothershberg"
date: "2022-05-26"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Notes for Unit 9: Principles of Regression Analysis

## Explanatory versus predictive modeling

- Different goals of regression modeling:
  1. Descriptive
  2. Predictive
  3. Explanatory

## Testing for confounders

- Rule of thumb: if a potential confounder changes the beta coefficient between the predictor of interest and the outcome variable by more than 10%, then it is considered a confounder.

## Residual confounding

- You cannot wipe out confounding simply by adjusting for variables in regression (there is always some measurement error).
- Residual confounding can plausibly generate spurious relative risks in the range of 0.6 to 1.6.
- Residual confounding can also obscure relationships

## Interactions

- Interaction is when the effect size is significantly different in different subgroups.
- Interactions are encoded as the product between terms in a regression model.
- Article: Open access increases citation impact for higher quality research
- "The effect is significant in 1 subgroup but not another" is not proof of interaction.

Misc: my current estimate of points before final:

```{r points}
# quizzes, homework, midterm 1, midterm 2
10 + (0.9 * 30) + (0.7 * 15) + (0.85 * 15)
```

Final is 30 points. So probably need 20/30 or greater.

## Regression worries

- Data can have nonlinear relationships
  - can add a quadratic term to models
- There can be influential data points that skew models
- P-hacking
  - tuning the form of the predictor
  - finding exact combination of confounders
  - "excluding" outliers
  - looking through subgroups
  - etc.
- Overfitting
  - rule of thumb: you need at least 10 subjects for each predictor variable in the regression model

## Unit 9 R lab

```{r unit-9-lab}
library(tidyverse)
library(readxl)

class_data <- read_excel("data/classdata.xlsx")
class_data$IsBookSmart <- as.factor(class_data$IsBookSmart)

class_data_filtered <- class_data %>%
  filter(!is.na(IsBookSmart))

ggplot(class_data_filtered, aes(x = politics, y = obama, color = IsBookSmart)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

model <- glm(obama ~ IsBookSmart + politics + IsBookSmart * politics,
             data = class_data)

summary(model)
```

## Homework

```{r unit-9-hw}
# 1
199.0439648 + (0.1239354 * 25) + -97.6187429 + (25 * 6.7372626)

# 13
class_data$varsity <- as.factor(class_data$varsity)
ggplot(class_data, aes(x = alcohol, y = optimism, color = varsity)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

model <- glm(optimism ~ alcohol + varsity + optimism * varsity,
             data = class_data)

summary(model)
```






