---
title: "unit-8"
author: "elliothershberg"
date: "2022-05-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Notes for Unit 8: Linear Regression

## Covariance and Correlation

- Covariance is how much two different traits vary together.

$$
Cov(x,y) = \frac{\sum_{i=1}^{n}(x_i - \bar X)(y_i - \bar Y)}{n - 1}
$$

- Relationship with variance: Var(x) = Cov(x, x)
- Correlation measures the strength of the linear relationship between two variables.
  - Ranges between -1 and 1
- Distribution of a correlation coefficient:
  - T-disributed with n = 2 degrees of freedom
  - mean = true correlation coefficient
  - standard error: $\sqrt{\frac{1 - r^2}{n - 2}}$
- Spearman's rank correlation uses ranks instead of raw variables

## Simple Linear Regression

- Predict response variable based on explanatory variables
- find line that best fits the data
- Expected value of y at a given level of x: $E(y_i | x_i) = \alpha + \beta x_i$
- predicted value: $\hat y_i = \alpha + \beta x_i + error_i$
- Assumptions:
  - relationship is linear
  - Y is distributed normally at each value of X
  - homogeneity of variances
  - observations are independent
- Least squares estimation is used to find the "best fit" line.
- Formulas:
  - Slope (beta coefficient): $\hat \beta = \frac{Cov(x,y)}{Var(x)}$
  - Intercept: $\hat \alpha = \bar y - \hat \beta \bar x$

## Residual Analysis

- residual = observed - predicted
- residual plots should show no pattern if the assumptions of regression are met.
- can check normality of residuals with a histogram.
- 
  
## Multiple linear regression

- Models with more than one predictor
- adjusting for confounding variable by adding it is a a predictor in the model

## Categorical predictors in regression

- treat binary variables as 0 or 1.
- A t-test is actually just linear regression.
  - Same story for ANOVA, etc.
  - This is useful because you can use regression to adjust for confounding factors.
- Dummy coding to create categorical variables.

## Unit 8 lab

```{r unit-8-lab}
library(readxl)
library(tidyverse)

classdata <- read_excel("data/classdata.xlsx")

ggplot(classdata, aes(x = bushjr, y = obama)) +
  geom_point() +
  geom_smooth()

cor.test(classdata$bushjr, classdata$obama)

model <- glm(obama ~ bushjr, data = classdata)

summary(model)

# plot(model)

hist(residuals(model))

shapiro.test(residuals(model))

model_2 <- glm(obama ~ bushjr + politics, data = classdata)
summary(model_2)
```

## Homework

```{r hw}
ggplot(classdata, aes(x = bushsr, y = bushjr)) +
  geom_point() +
  geom_smooth()

cor.test(classdata$bushsr, classdata$bushjr)

model <- glm(bushjr ~ bushsr, data = classdata)
summary(model)

model_2 <- glm(bushjr ~ bushsr + politics, data = classdata)
summary(model_2)
```




