---
title: "module-1"
author: "elliothershberg"
date: "2022-04-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Notes from Unit 1: Descriptive Statistics and Looking at Data

### Types of data (in medicine)

-   Quantitative
-   Categorical (nominal or ordinal)
    -   binary
    -   nomimal (unordered categories)
    -   ordinal (ordered categories)
-   Time-to-event
    -   the time it takes for an event to occur (if it occurs at all)
    -   binary part (yes/no)
    -   continuous part (time)

### Visualizing data

-   Always plot your data
    -   identify errors
    -   see distributions
-   Frequency plots (univariate)
    -   bar charts
    -   histograms
    -   boxplots (how to draw whiskers and outliers)
-   Shapes of distributions
    -   normal distribution

### Describing quantitative data: where is the center?

-   Measures of central tendency
    -   mean
        -   it is like a balancing point. The farther away from the mean, the more a value will affect the mean.
        -   binary variables have means
    -   median (middle value of ordered data)
    -   Overview of Wilcoxon Rank-Sum Test
        -   Rank two sets of values
        -   sum the ranks for each set
        -   compare

### Describing quantitative data: standard deviation, percentiles

-   range
-   standard deviation/variance
    -   most common measure of variability in statistics
    -   variance is the average squared distance from the mean
    -   standard deviation is the square root of the variance
    -   standard error measures the variability of a statistic
-   percentiles
-   interquartile range
-   degrees of freedom: the number of values in the final calculation of a statistic that are free to vary.

$$
S^2 = \frac{\sum(x_i - \bar x)^2}{n - 1}
$$

### Homework

```{r sleep-data}
sleep_data <- rep(c(5, 6, 7, 8, 9, 10), times = c(3, 8, 20, 11, 6, 2))
```


1.  

```{r problem-1}
sleep_sum <- 5*3 + 6*8 + 7*20 + 8*11 + 9*6 + 10*2

sleep_mean <- sleep_sum / 50

sleep_mean
```

2.  

```{r problem-2}
squared_dist <- function(x, m = sleep_mean) {
  (x - m)^2
}

sleep_variance <- (3*squared_dist(5) + 8*squared_dist(6) + 20*squared_dist(7) + 11*squared_dist(8) + 6*squared_dist(9) + 2*squared_dist(10)) / 49

standard_dev <- sqrt(sleep_variance)

standard_dev
```

7.

```{r problem-7}
coin_mean <- (0*47 + 1*53) / 100

coin_mean
```

8.

```{r problem-8}
coin_vec <- rep(c(0, 1), times = c(47, 53))

sd(coin_vec)
```

I need to load class data for the last 3 problems.

```{r load-class-data}
library(readxl)

class_data <- read_excel("data/classdata.xlsx")
```


17.

```{r problem-17}
sd(class_data$fruitveg)
```

18.

```{r problem-18}
median(class_data$fruitveg)
```

19.

```{r problem-19}
hist(class_data$coffee, breaks = 7)
```

